### 2020-02-12 reports/raw_data_viz.Rmd
Add report visualizing raw data

---
title: "Data plots"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load packages and custom functions
source(here::here("code/setup.R"))
source(here::here("code/functions.R"))
```

Visualize the data and check for any outliers.

## Climate

Load raw data.

```{r load-climate-data}
# Read in raw climate data
climate_raw <- read_csv(here::here("data/hobo_moorea_aorai_2-24-15.csv"))

# Subset to Moorea, format site as 
climate_raw_moorea <- climate_raw %>%
   filter(str_detect(site, "Aorai", negate = TRUE)) %>%
   mutate(date_time = (lubridate::ymd(date) + lubridate::hms(time)) %>% as.POSIXct())
```

Identify time periods with erroneous values to exclude from cleaned dataset. These are the values that will be excluded by `process_raw_climate()`.

Note the band of red from 2013-11-19 to 2014-03-21 is actually OK for most dataloggers, but not for Tohiea 1170m ter + Tohiea 2000m ter, so it is flagged to be excluded from all sites.

```{r format-raw-climate}
# Make dataframe of dates/sites to flag as having weird data.
highlight_chunks <- 
  climate_raw_moorea %>%
  select(date, site) %>%
  unique %>%
  mutate(flag = case_when(
    date >= "2014-03-12" & date < "2014-07-06" & site == "Tohiea_400m_ter" ~ TRUE,
    site == "Rotui_600m_epi" ~ TRUE,
    site == "Mouaputa_800m_epi" ~ TRUE,
    date >= "2013-11-19" & date <= "2014-03-21" ~ TRUE,
    date >= "2013-07-26" & date <= "2013-07-29" & site == "Tohiea_600m_epi"~ TRUE,
    date == "2013-08-08" & site == "Tohiea_600m_ter" ~ TRUE,
    date %in% c("2014-06-15", "2014-06-16", "2014-06-19", "2014-06-20") & site == "Tohiea_600m_ter" ~ TRUE,
    TRUE ~ FALSE
  ))

# Add the "flag" column back into the original data
climate_raw_moorea_highlight <-
  left_join(climate_raw_moorea, highlight_chunks, by = c("site", "date")) %>%
  # Reformat growth habit as factor, separate from site name
   mutate(
      habit = case_when(
        str_detect(site, "epi") ~ "epiphytic",
        str_detect(site,  "ter") ~ "terrestrial"
      ),
      habit = fct_relevel(habit, c("epiphytic", "terrestrial")),
      site = str_remove_all(site, "_epi|_ter")
    )
```

Plot the raw data, highlighting the data that will be excluded from the cleaned dataset in red.

```{r raw-temp-plot, fig.height = 24, fig.width = 8}
ggplot(climate_raw_moorea_highlight, aes(x = date_time, y = temp, color = flag)) +
  geom_line() +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  labs(
    title = "Moorea temperature",
    subtitle = "raw data") +
  theme(legend.position = "bottom") +
  facet_grid(vars(site), vars(habit))
```

```{r raw-rh-plot, fig.height = 24, fig.width = 8}
ggplot(climate_raw_moorea_highlight, aes(x = date_time, y = rh, color = flag)) +
  geom_line() +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  labs(
    title = "Moorea rel. humidity",
    subtitle = "raw data") +
  theme(legend.position = "bottom") +
  facet_grid(vars(site), vars(habit))
```

Process raw data, excluding time periods with erroneous values.

```{r process-climate}
# Process raw climate data: 
# subset to only Moorea and days without missing data,
# replace one chunk of missing data with same dates from previous year,
# calculate vapor pressure deficit (VPD) from temp and RH.
moorea_climate <- process_raw_climate(climate_raw) %>%
  # Add date_time column for plotting
   mutate(date_time = (lubridate::ymd(date) + lubridate::hms(time)) %>% as.POSIXct())
```

Plot the cleaned climate data.

```{r clean-temp-plot, fig.height = 24, fig.width = 8}
ggplot(moorea_climate, aes(x = date_time, y = temp)) +
  geom_line() +
  labs(
    title = "Moorea temperature",
    subtitle = "cleaned data") +
  facet_grid(vars(site), vars(habit))
```

```{r clean-rh-plot, fig.height = 24, fig.width = 8}
ggplot(moorea_climate, aes(x = date_time, y = rh)) +
  geom_line() +
  labs(
    title = "Moorea rel. humidity",
    subtitle = "cleaned data") +
  facet_grid(vars(site), vars(habit))
```