### 2020-02-13 reports/raw_data_viz.Rmd
Update raw data viz report

---
title: "Data plots"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
# Load packages and custom functions
source(here::here("code/setup.R"))
source(here::here("code/functions.R"))
```

Visualize the data and check for any outliers.

## Climate

Load raw data.

```{r load-climate-data}
# Read in raw climate data
climate_raw <- read_csv(here::here("data/hobo_moorea_aorai_2-24-15.csv"))

# Subset to Moorea
climate <- climate_raw %>%
    filter(str_detect(site, "Aorai", negate = TRUE)) %>%
    mutate(
      day = lubridate::day(date) %>% as.numeric,
      month = lubridate::month(date) %>% as.numeric,
      year = lubridate::year(date) %>% as.numeric,
      # add date_time col for plotting
      date_time = (lubridate::ymd(date) + lubridate::hms(time)) %>% as.POSIXct())
```

Identify time periods with erroneous values to exclude from cleaned dataset. Do this with code first below and inspect the results.

```{r flag-extremes}
# Identify start and end dates of survey including all sites on Moorea
survey_window <-
climate %>%
  group_by(site) %>%
  summarize(
    last_measured_day = max(date),
    first_measured_day = min(date)
  ) %>%
  summarize(
    end_date = min(last_measured_day),
    start_date = max(first_measured_day)
  ) %>%
  mutate_all(as.POSIXct)

# Tally days with extreme (possibly erroneous) data
extremes_tally <-
climate %>%
  # Exclude two sites completely that are mostly missing data
  filter(site != "Rotui_600m_epi") %>%
  filter(site != "Mouaputa_800m_epi") %>%
  # Crop start and end dates
  filter(date > as.Date(survey_window$start_date) & date < as.Date(survey_window$end_date)) %>%
  # Flag any days as possible bad data if values are extremely low or high
  mutate(
    temp_hi = ifelse(temp > 50, TRUE, FALSE),
    temp_low = ifelse(temp < 5, TRUE, FALSE),
    rh_low = ifelse(rh < 10, TRUE, FALSE)) %>%
  group_by(site, date) %>%
  summarize(
    temp_hi = sum(temp_hi),
    temp_low = sum(temp_low),
    rh_low = sum(rh_low)
  ) %>%
  ungroup %>%
  filter_at(vars(temp_hi, temp_low, rh_low), any_vars(. > 0)) %>%
  mutate(flag = TRUE) %>%
  arrange(site, date)

extremes_tally %>% DT::datatable()
```

After inspecting raw data plots and the points identified above, flag days for removal by hard-coding.

```{r format-raw-climate}
# Make dataframe of dates/sites to flag as having weird data.
climate_data_flags <- 
  climate %>%
  select(date, site) %>%
  unique %>%
  mutate(flag_rh = case_when(
    # Flag 2013-11-19 to 2013-12-15 (Tohiea 1170m ter 0% RH)
    date >= "2013-11-19" & date <= "2013-12-15" & site == "Tohiea_1170m_ter" ~ TRUE,
    # Flag 2013-12-29 to 2014-03-12 (Tohiea 200m ter unusually low RH)
    date >= "2013-12-29" & date <= "2014-03-12" & site == "Tohiea_200m_ter" ~ TRUE,
    # Flag 2014-06-15 to 2014-06-20 (Tohiea_800m_ter 0% RH)
    date >= "2014-06-15" & date <= "2014-06-20" & site == "Tohiea_800m_ter" ~ TRUE,
    # Flag 2013-08-08 (Tohiea_600m_ter 1% RH)
    date == "2013-08-08" & site == "Tohiea_600m_ter" ~ TRUE,
    # Flag Tohiea sites missing data
    date >= "2014-03-13" & date <= "2014-03-20" & str_detect(site, "Tohiea") ~ TRUE,
    TRUE ~ FALSE
  )) %>%
  mutate(flag_temp = case_when(
    # Delete 2013-07-26 to 2013-07-29 (Tohiea_600m epi temperature below 0C)
    date >= "2013-07-26" & date <= "2013-07-29" & site == "Tohiea_600m_epi" ~ TRUE,
    # Flag Tohiea sites missing data
    date >= "2014-03-13" & date <= "2014-03-20" & str_detect(site, "Tohiea") ~ TRUE,
    TRUE ~ FALSE
  ))

# Add the "flag" column back into the original data
climate_flagged <- left_join(climate, climate_data_flags, by = c("site", "date"))
```

Plot the raw data, highlighting the data that will be excluded from the cleaned dataset in red, and lines indicating cropped starting and ending dates in blue.

```{r raw-temp-plot, fig.height = 24, fig.width = 8}
ggplot(climate_flagged, aes(x = date_time, y = temp, color = flag_temp)) +
  geom_line() +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  annotate("segment", 
           x = survey_window$start_date,
           xend = survey_window$start_date, 
           y = -Inf, yend = Inf, color = "blue") +
  annotate("segment", 
           x = survey_window$end_date, 
           xend = survey_window$end_date, 
           y = -Inf, yend = Inf, color = "blue") +
  labs(
    title = "Moorea temperature",
    subtitle = "raw data") +
  theme(legend.position = "bottom") +
  facet_wrap(vars(site), ncol = 2)
```

```{r raw-rh-plot, fig.height = 24, fig.width = 8}
ggplot(climate_flagged, aes(x = date_time, y = rh, color = flag_rh)) +
  geom_line() +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black")) +
  annotate("segment", 
           x = survey_window$start_date,
           xend = survey_window$start_date, 
           y = -Inf, yend = Inf, color = "blue") +
  annotate("segment", 
           x = survey_window$end_date, 
           xend = survey_window$end_date, 
           y = -Inf, yend = Inf, color = "blue") +
  labs(
    title = "Moorea rel. humidity",
    subtitle = "raw data") +
  theme(legend.position = "bottom") +
  facet_wrap(vars(site), ncol = 2)
```

Process raw data, excluding time periods with erroneous values. Any day with erroneous values for any site will be excluded from ALL sites.

```{r process-climate}
# Process climate using same code as function in analysis:

# Subset to Moorea
climate_moorea <- climate_raw %>%
    filter(str_detect(site, "Aorai", negate = TRUE)) %>%
    mutate(
      day = lubridate::day(date) %>% as.numeric,
      month = lubridate::month(date) %>% as.numeric,
      year = lubridate::year(date) %>% as.numeric)

# Split out set of days from 2013 and 2014 to use for filling in 
# missing 2014 Tohiea 400 m ter data
tohiea_400m_2013 <- climate_moorea %>%
  filter(date >= "2013-03-12" & date < "2013-07-06" & site == "Tohiea_400m_ter")

tohiea_400m_2014 <- climate_moorea %>%
  filter(date >= "2014-03-12" & date < "2014-07-06" & site == "Tohiea_400m_ter")

# Replace the year 2013 with 2014 only for these dates in the Tohiea 400 m ter data
tohiea_400m_as_2014 <-
  tohiea_400m_2013 %>%
  mutate(year = year + 1) %>%
  mutate(date = paste(year, month, day, sep = "-") %>% as.Date)

# Process data
climate_filtered <- climate_moorea %>%
  # Remove bad Tohiea data from 2014
  anti_join(tohiea_400m_2014, by = c("date", "site")) %>%
  # Replace bad Tohiea data from 2014 with data from 2013
  bind_rows(tohiea_400m_as_2014) %>%
  # Remove two sites that have are (nearly) complete failures
  filter(site != "Rotui_600m_epi") %>%
  filter(site != "Mouaputa_800m_epi") %>%
  # Crop to dates when all dataloggers were set up on Moorea
  filter(date >= "2013-07-07" & date <= "2014-07-05") %>%
  # Flag dates to remove:
  mutate(
    flag = case_when(
      # 2013-11-19 to 2013-12-15 (Tohiea 1170m ter 0% RH)
      date >= "2013-11-19" & date <= "2013-12-15" ~ TRUE,
      # 2013-12-29 to 2014-03-12 (Tohiea 200m ter unusually low RH)
      date >= "2013-12-29" & date <= "2014-03-12" ~ TRUE,
      # 2014-03-13 to 2014-03-20 (Missing data for most Tohiea sites)
      date >= "2014-03-13" & date <= "2014-03-20" ~ TRUE,
      # 2014-06-15 to 2014-06-20 (Tohiea_800m_ter 0% RH)
      date >= "2014-06-15" & date <= "2014-06-20" ~ TRUE,
      # 2013-07-26 to 2013-07-29 (Tohiea_600m epi temperature below 0C)
      date >= "2013-07-26" & date <= "2013-07-29" ~ TRUE,
      # 2013-08-08 (Tohiea_600m_ter 1% RH)
      date == "2013-08-08" ~ TRUE,
      TRUE ~ FALSE
    )) %>%
  # Instead of filtering out flagged data, set temp and RH for these to NA
  # this will make them appear as gaps in the plot
  mutate(
    rh = case_when(
      flag == FALSE ~ rh,
      flag == TRUE ~ NaN
    )
  ) %>%
  mutate(
    temp = case_when(
      flag == FALSE ~ temp,
      flag == TRUE ~ NaN
    )
  ) %>%
  select(-flag) %>%
  # Check that there are no duplicate measurements for a given time/site/date
  assert_rows(col_concat, is_uniq, date, time, site) %>%
  # Check for temp, RH in reasonable range 
  # (shouldn't be any freezing, expect at least some humidity)
  assert(within_bounds(0,50, include.lower = FALSE), temp) %>%
  assert(within_bounds(1,100), rh) %>%
  # Reformat filtered data, add VPD
  mutate(
    habit = case_when(
      str_detect(site, "epi") ~ "epiphytic",
      str_detect(site,  "ter") ~ "terrestrial"
    ),
    habit = fct_relevel(habit, c("epiphytic", "terrestrial")),
    site = str_remove_all(site, "_epi|_ter"),
    vpd = plantecophys::RHtoVPD(rh, temp),
    # add date_time col for plotting
    date_time = (lubridate::ymd(date) + lubridate::hms(time)) %>% as.POSIXct()
  )
```

Plot the cleaned climate data.

```{r clean-temp-plot, fig.height = 24, fig.width = 8}
ggplot(climate_filtered, aes(x = date_time, y = temp)) +
  geom_line() +
  labs(
    title = "Moorea temperature",
    subtitle = "cleaned data") +
  facet_grid(vars(site), vars(habit))
```

```{r clean-rh-plot, fig.height = 24, fig.width = 8}
ggplot(climate_filtered, aes(x = date_time, y = rh)) +
  geom_line() +
  labs(
    title = "Moorea rel. humidity",
    subtitle = "cleaned data") +
  facet_grid(vars(site), vars(habit))
```

## Traits

Load raw data and prepare for plotting.

```{r load-trait-data}
# Read in species list
species_list <- read_csv(here::here("data/nitta_2017/species.csv")) %>%
    clean_names() %>%
    filter(include == 1, tahiti_only == 0) %>%
    # Exclude Microsorum_xmaximum (hybrid between M. grossum and M. commutatum)
    filter(genus_sp != "Microsorum_xmaximum") %>%
    pull(genus_sp)

# Process raw continuous trait data
# (includes one measure per specimen,
#  multiple specimens per species).
# All measurements in cm, except number of pinna pairs.
  morph_cont_raw <- process_raw_cont_morph(
    here::here("data/morph_measurements_2020-02-10.csv"),
    species_list
  )

# Angiopteris evecta is much bigger than other species, obscures
# points in plot. Exclude this outlier for plotting.
morph_cont_raw_no_evecta <-
  morph_cont_raw %>%
  filter(!str_detect(species, "evecta"))

# Define a function for plotting trait data

plot_raw_quant_morph <- function(morph_cont_raw, trait_select) {

  # Subset to selected trait
  trait_select_enquo <- enquo(trait_select)

  morph_cont_sel <- select(morph_cont_raw, species, !!trait_select_enquo)

  # Calculate mean trait value to arrange species box plots by mean
  means <-
    morph_cont_sel %>%
    group_by(species) %>%
    summarize(
      mean = mean(!!trait_select_enquo, na.rm = TRUE)
    )

  morph_cont_sel %>%
    left_join(means) %>%
    mutate(species = as.factor(species) %>% fct_reorder(mean)) %>%
    ggplot(aes(x = species, y = !!trait_select_enquo)) +
    geom_boxplot() +
    coord_flip()

}
```

Plot raw quantitative trait data.

```{r plot-stipe, fig.height = 16, fig.width = 6}
plot_raw_quant_morph(morph_cont_raw_no_evecta, stipe_length) +
  labs(title = "Stipe length")
```

```{r plot-frond, fig.height = 16, fig.width = 6}
plot_raw_quant_morph(morph_cont_raw_no_evecta, frond_length) +
  labs(title = "Frond length")
```

```{r plot-lamina-width, fig.height = 16, fig.width = 6}
plot_raw_quant_morph(morph_cont_raw_no_evecta, lamina_width) +
  labs(title = "Lamina width")
```

```{r plot-rhizome, fig.height = 16, fig.width = 6}
plot_raw_quant_morph(morph_cont_raw_no_evecta, rhizome_dia) +
  labs(title = "Rhizome diameter")
```

```{r plot-pinna-pairs, fig.height = 16, fig.width = 6}
plot_raw_quant_morph(morph_cont_raw_no_evecta, pinna_pairs) +
  labs(title = "No. pinna pairs")
```