### 2020-02-19 ms/SI.rmd
Improve table S2

---
title: ""
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
- BoldFont=Roboto-Bold.ttf
- ItalicFont=Roboto-Italic.ttf
- BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: none
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "new-phytologist.csl"]
    includes:
      in_header: new-phytologist.sty
header-includes: 
  - \usepackage{float}
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother # Fix position of figs
  - \makeatletter\renewcommand*{\fps@table}{H}\makeatother # Fix position of tables
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE)

# Load analysis results
loadd(list = c(
  "taxonomic_data_for_si",
  "trait_data_for_si",
  "climate_ancova_results",
  "phylosig_binary_traits_strict",
  "correlated_evo_test_strict",
  "best_fit_div_models_moran",
  "func_div_corr_plot",
  "climate_model_summaries",
  "climate_models",
  "div_el_models",
  "div_el_model_summaries",
  "div_el_model_parameter_summaries",
  "cwsd_scatterplot"), 
  cache = canopy_cache)

# Load captions
source(here::here("ms/captions.R"))

# Load functions
source(here::here("code/functions.R"))

# Set up main table header
si_table_header <- c(
  "New Phytologist Supporting Information",
  "Authors: Joel H. Nitta, James E. Watkins, Jr., and Charles C. Davis" %>% wrap_quotes,
  "Article acceptance date: 5 February 2020"
)

```

<!-- Set left justification  -->
\raggedright

<!-- Don't indent -->
\setlength{\parindent}{0ex}

***New Phytologist*** **Supporting Information**

Article title: Life in the canopy: Community trait assessments reveal substantial functional diversity among fern epiphytes

Authors: Joel H. Nitta, James E. Watkins, Jr., and Charles C. Davis

Article acceptance date: 5 February 2020

The following Supporting Information is available for this article:

**`r s_table("species-list")`** `r s_table_cap("species-list")` (see separate file)

```{r species-list}
species_list_footer <- "'Taxon code' is the name used for that taxon in analyses. 'Informal variety' is a number assigned to infraspecific taxa that are morphologically and genetically distinct, but lack formal names distinguishing them. 'Name source' indicates the source of the scientific name as follows: IPNI = International Plant Names Index (https://www.ipni.org/), TROPICOS = Missouri Botanical Garden taxonomic database (https://www.tropicos.org/), NCBI = National Center for Biotechnology Information taxonomic database (https://www.ncbi.nlm.nih.gov/Taxonomy/Browser/wwwtax.cgi), Murdock and Smith 2003 = Murdock & Smith (2003). The genus in 'Taxon code' may differ from the 'Genus' column for names whose preferred synonymy has changed since data were originally recorded. 'NA' indicates not applicable."

write_si_table(
  table_id = "species-list", 
  main_header = si_table_header,
  table = taxonomic_data_for_si,
  table_footer = species_list_footer
)
```

`r species_list_footer`

**`r s_table("trait-values")`** `r s_table_cap("trait-values")` (see separate file)

```{r trait_data}
trait_values_footer <- "SLA, specific leaf area; 'Taxon code', name used for that taxon in analysis (see Table S1); 'Growth habit', growth habit observed in the field; 'Binary growth habit', binary growth habit used in analysis (see Methods). Continuous measurements shown as mean, standard deviation (sample size). For data source, L = lab observation of gametophyte cultured from spore, F = field observation of gametophyte subsequently identified using DNA barcoding, T = morphology inferred based on taxonomy, M = measurement of a field-collected specimen; cited references provided at end of SI. For frond dissection, 1 = simple, 2 = pinnatifid, 3 = one-pinnate, 4 = one-pinnate-pinnatifid, 5 = two-pinnate, 6 = two-pinnate-pinnatifid, 7 = bipinnatifid_or_tripinnatifid, 8 = three-pinnate, 9 = three-pinnate-pinnatifid. 'NA' indicates missing data."

# Format table: capitalize column names
trait_values_table <-
trait_data_for_si %>% 
  rename_all(str_to_sentence) %>% 
  rename(`SLA (sq m per kg)` = `Sla (sq m per kg)`)

write_si_table(
  table_id = "trait-values", 
  table = trait_values_table,
  main_header = si_table_header,
  table_footer = trait_values_footer
)
```

`r trait_values_footer`

**`r s_table("ancova")`** `r s_table_cap("ancova")` (see separate file)

```{r ancova}
# Format table
ancova_table <-
  tribble(
  ~variable, ~Response,
  "temp_max", "Max. temp.",
  "temp_mean", "Mean temp.",
  "temp_sd", "SD temp.",
  "vpd_min", "Min. VPD",
  "vpd_mean", "Mean VPD",
  "vpd_sd", "SD VPD",
) %>%
  left_join(climate_ancova_results, by = "variable") %>% 
  mutate(
    p.value = signif(p.value, 2) %>% format.pval(., digits = 1),
    meansq = round_t(meansq, 2),
    term = case_when(
      term == "habit" ~ "Growth habit",
      term == "el" ~ "Elevation",
      term == "habit:el" ~ "Growth habit × elevation",
      TRUE ~ term
    )
  ) %>%
  select(Response, `Source of variation` = term, `d.f.` = df, MS = meansq, P = p.value)

ancova_footer <- "d.f., degrees of freedom; MS, mean squares. Response variables include mean, minimum, maximum, and standard deviation of temperature ('temp', °C) and vapor pressure deficit ('VPD', kPa). Vapor pressure deficit was square-root transformed prior to analysis. 'NA' indicates not applicable."

write_si_table(
  table_id = "ancova", 
  table = ancova_table,
  main_header = si_table_header,
  table_footer = ancova_footer
)
```

`r ancova_footer`

**`r s_table("linear-fit")`** `r s_table_cap("linear-fit")` (see separate file)

```{r linear-fit}
# Combine summaries of models for diversity metrics
div_summaries <-
  div_el_model_summaries %>%
  left_join(
    select(div_el_models,
           var, model_type, delta_AICc, morans_I, I_pval), 
    by = c("var", "model_type")
  )

# Combine summaries of models for climate metrics
climate_summaries <-
  climate_model_summaries %>%
  left_join(
    select(climate_models,
           var, model_type, delta_AICc, morans_I, I_pval), 
    by = c("var", "model_type")
  )

# Combine summaries of diversity and climate metrics, reformat model names
combined_summaries <-
  bind_rows(div_summaries, climate_summaries) %>%
  mutate(
    model_type = model_type %>% 
      str_replace_all("el_only", "Elevation") %>%
      str_replace_all("both", "Elevation + Growth habit") %>%
      str_replace_all("habit_only", "Growth habit") %>%
      str_replace_all("interaction", "Elevation + Growth habit + Elevation x Growth habit")
    )

# Reformat variable names and put in preferred order by joining
linear_fit_table <-
tribble(
  ~var, ~Response,
  "temp_max", "Max. temp.",
  "temp_mean", "Mean temp.",
  "temp_sd", "SD temp.",
  "vpd_min", "Min. VPD",
  "vpd_mean", "Mean VPD",
  "vpd_sd", "SD VPD",
  "ntaxa", "Richness",
  "mpd.obs.z.phy", "MPDphy",
  "mntd.obs.z.phy", "MNTDphy",
  "mpd.obs.z.func", "MPDfunc",
  "mntd.obs.z.func", "MNTDfunc",
  "dissection", "Dissection",
  "length", "Frond length",
  "pinna", "Pinna no.",
  "rhizome", "Rhizome diam.",
  "stipe", "Stipe length",
  "sla", "SLA",
  "width", "Frond width"
) %>%
  left_join(combined_summaries, by = "var") %>%
  select(
    Response,
    Model = model_type,
    AICc,
    f = statistic,
    `Adjusted R2` = adj.r.squared,
    P = p.value,
    `d.f.` = df,
    `delta AICc` = delta_AICc,
    `Moran's I` = morans_I,
    `P moran` = I_pval
  )

linear_fit_footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; VPD, vapor pressure deficit; SLA, specific leaf area; AICc, corrected Akaike Information Criterion; d.f., degrees of freedom; 'P moran', P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I. Trait names refer to community-weighted mean values (see Methods). For details on calculations of MPD and MNTD, see Methods. 'NA' indicates not applicable."

write_si_table(
  table_id = "linear-fit", 
  table = linear_fit_table,
  main_header = si_table_header,
  table_footer = linear_fit_footer
)
```

`r linear_fit_footer`

**`r s_table("phylosig-binary-strict")`** `r s_table_cap("phylosig-binary-strict")` (see separate file)

```{r phylosig-binary-strict}
# Format table
phylosig_binary_strict_table <-
  phylosig_binary_traits_strict %>%
  mutate(D = round_t(D, 2)) %>%
  mutate(trait = factor(
    trait, 
    levels = c("habit", "gemmae", "glands", "hairs","morphotype"))) %>%
  mutate(trait = lvls_revalue(
    trait,
    new_levels = c("Epiphytic growth", "Gemmae", "Glands", "Hairs", "Cordate morphotype"))) %>%
  arrange(trait) %>%
  select(Trait = trait, 
         `Number of presences` = num_present, 
         `Number of absences` = num_absent, 
         D,
         `P rnd` = prob_random, 
         `P bm` = prob_brownian)

phylosig_binary_strict_footer <- "'P rnd', probability of obtaining the observed value of D under random trait evolution; 'P bm', probability of obtaining the observed value of D under simulation of trait evolution by Brownian motion."

write_si_table(
  table_id = "phylosig-binary-strict", 
  table = phylosig_binary_strict_table,
  main_header = si_table_header,
  table_footer = phylosig_binary_strict_footer
)
```

'P rnd', probability of obtaining the observed value of \D under random trait evolution; 'P bm', probability of obtaining the observed value of \D under simulation of trait evolution by Brownian motion.

**`r s_table("corr-evo-strict")`** `r s_table_cap("corr-evo-strict")` (see separate file)

```{r corr-evo-strict}
# Format table
corr_evo_strict_table <-
  correlated_evo_test_strict %>%
  mutate(trait = factor(
    trait, 
    levels = c("gemmae", "glands", "hairs","morphotype"))) %>%
  mutate(trait = lvls_revalue(
    trait,
    new_levels = c("Gemmae", "Glands", "Hairs", "Morphotype"))) %>%
  arrange(trait) %>%
  select(
    Trait = trait,
    `LL (independent model)` = logL_indep,
    `LL (dependent model)` = logL_dep,
    `Likelihood ratio` = likelihood_ratio,
    `P` = pval
  )

corr_evo_strict_footer <- "LL, log likelihood. Each gametophyte trait was coded as a binary trait and tested for correlated evolution with growth habit (epiphytic vs. terrestrial). Lower log likelihood for the dependent model indicates correlation with growth habit."

write_si_table(
  table_id = "corr-evo-strict", 
  table = corr_evo_strict_table,
  main_header = si_table_header,
  table_footer = corr_evo_strict_footer
)
```

`r corr_evo_strict_footer`

**`r s_table("gam-fit")`** `r s_table_cap("gam-fit")` (see separate file)

```{r gam-fit}
# Format table

# First make join table for converting variable names, in order
resp_var_table <-
tribble(
  ~resp_var, ~Response,
  "ntaxa", "Richness",
  "mpd.obs.z.phy", "MPDphy",
  "mntd.obs.z.phy", "MNTDphy",
  "mpd.obs.z.func", "MPDfunc",
  "mntd.obs.z.func", "MNTDfunc",
  "dissection", "Dissection",
  "length", "Frond length",
  "width", "Frond width",
  "pinna", "Pinna no.",
  "rhizome", "Rhizome diam.",
  "stipe", "Stipe length",
  "sla", "SLA"
)

gam_fit_table <-
resp_var_table %>%
  left_join(best_fit_div_models_moran, by = "resp_var") %>%
  mutate(resp_var = fct_relevel(resp_var, resp_var_table$resp_var)) %>%
  arrange(resp_var, delta_AICc) %>%
  mutate(
    modname = modname %>%
      str_replace_all("\\+", " \\+ ") %>%
      str_replace_all("\\.by\\.", " x ") %>%
      str_replace_all("habit", "Growth habit") %>%
      str_replace_all("temp_mean", "Mean temp.") %>%
      str_replace_all("temp_sd", "SD temp.") %>%
      str_replace_all("temp_max", "Max. temp.") %>%
      str_replace_all("vpd_mean", "Mean VPD") %>%
      str_replace_all("vpd_sd", "SD VPD") %>%
      str_replace_all("vpd_min", "Min. VPD")
  ) %>% 
  select(
    Response,
    `Model` = modname,
    AICc,
    `R2` = r2,
    `delta AICc` = delta_AICc,
    `Moran's I` = morans_I,
    `P moran` = morans_I_pval
  )

gam_fit_footer <- "MPDphy, standard effect size of mean phylogenetic distance; MNTDphy, standard effect size of mean nearest taxon distance; MPDfunc, standard effect size of mean functional distance; MNTDfunc, standard effect size of mean nearest functional distance; SLA, specific leaf area; VPD, vapor pressure deficit, AICc, corrected Akaike Information Criterion; 'P moran', P value for the null hypothesis that there is spatial autocorrelation in the residuals as measured with Moran's I. Models within delta AICc 3 of the best-fitting model shown. Trait names refer to community-weighted mean values (see Methods). For details on calculations of MPD and MNTD, see Methods."

write_si_table(
  table_id = "gam-fit", 
  table = gam_fit_table,
  main_header = si_table_header,
  table_footer = gam_fit_footer
)

```

`r gam_fit_footer`

**`r s_table("spatial-sla")`** `r s_table_cap("spatial-sla")` (see separate file)

```{r spatial-sla}
# Format table
spatial_sla_table <-
div_el_model_parameter_summaries %>%
  filter(var == "sla") %>%
  select(term, Coefficient = estimate, SE = std.error, t = t_value) %>%
  mutate(term = term %>%
           str_replace_all("habitterrestrial", "Growth habit") %>%
           str_replace_all("el", "Elevation")) %>%
  rename(Term = term)

spatial_sla_footer <- "A Mátern spatial correlation structure was included in the model to account for spatial autocorrelation (see Methods)."

write_si_table(
  table_id = "spatial-sla", 
  table = spatial_sla_table,
  main_header = si_table_header,
  table_footer = spatial_sla_footer
)

```

`r spatial_sla_footer`

\clearpage

```{r Fig-S1, fig.height = 6, fig.width = 7, fig.cap = ""}
cwsd_scatterplot
```

**`r s_figure("cwsd")`** `r s_figure_cap("cwsd")`. Epiphytic communities in green, terrestrial communities in brown. Points randomly jittered along x-axis to reduce overlap. Boxplots show median values (bold lines) by growth habit. Lower and upper hinges correspond to first and third quartiles, and whiskers extend to values within 1.5 × the interquartile range. Asterisks indicate significant differences in CWSD between epiphytic and terrestrial communities as determined by a two-sided \tval-test; \*\*\* = \pval{ }< 0.001; \* = \pval{ }< 0.05 **(a)**, traits related to plant size **(b)**, traits not related to plant size. Response variable abbreviations: SLA, specific leaf area.

\clearpage

```{r Fig-S2, fig.height = 8, fig.width = 4, fig.cap = ""}
func_div_corr_plot
```

**`r s_figure("corr-plot")`** `r s_figure_cap("corr-plot")`. **(a)**, standard effect size of mean phylogenetic distance (MPD~func~) **(b)**, standard effect size of mean nearest taxon distance (MNTD~func~). For details on calculations of MPD~func~ and MNTD~func~, see Methods. Correlations between functional diversity as measured with different sets of traits (values in each tile) calculated with Pearson's correlation coefficient using the R package 'corrr'. All trait sets were scaled to standard deviation of each trait. The following trait sets were used: 'all' (all traits), 'size' (sporophyte traits related to size; frond length and width, stipe length, rhizome width), 'other' (sporophyte traits not related to size; pinna number, SLA, degree of lamina dissection), 'gameto' (gametophyte traits; gametophyte morphotype, presence or absence of glands, hairs, and gemmae), 'reduced' (all traits except for frond length and width, which were highly correlated with stipe length and rhizome diameter). For trait sets that included traits with skewed distributions in their raw values (frond length and width, stipe length, and rhizome width), both log-transformed and untransformed sets were included. The log-transformed 'reduced' dataset was used for the main results.

\clearpage

**References**

<!-- There is a bug somewhere in Rmd -> tex -> pdf using pandoc such that the 
first bold words until a period in each reference do not appear bold. As a 
work-around, first render to pdf with `bibliography` and `nocite` in the yaml, 
then convert Rmd -> tex -> docx with pandoc, edit docx manually to fix formatting, 
then docx -> md with pandoc, and paste properly formatted refs below -->

**Atkinson L**. **1975**. The gametophyte of five Old World thelypteroid
ferns. *Phytomorphology* **25**: 38--54.

**Bierhorst DW**. **1967**. The gametophyte of *Schizaea dichotoma*.
*American Journal of Botany* **54**: 538--549.

**Brownsey PJ**. **1987**. A review of the fern genus *Hypolepis*
(Dennstaedtiaceae) in the Malesian and Pacific regions. *Blumea* **32**:
227--276.

**Copeland EB**. **1932**. Pteridophytes of the Society Islands.
*Bernice P. Bishop Museum Bulletin* **93**: 1--86.

**Lloyd RM**. **1980**. Reproductive biology and gametophyte morphology
of New World populations of *Acrostichum aureum*. *American Fern
Journal* **70**: 99--110.

**Martin K, Sini S, Zhang C, Slater A, Madhusoodanan P**. **2006**.
Efficient induction of apospory and apogamy in vitro in silver fern
(*Pityrogramma calomelanos* L.). *Plant Cell Reports* **25**:
1300--1307.

**Murdock AG, Smith AR**. **2003**. Pteridophytes of Moorea, French
Polynesia, with a new species, *Tmesipteris gracilis* (Psilotaceae).
*Pacific Science* **57**: 253--265.

**National Museum of Nature and Science**. **2008**. *Illustrated Flora
of Ferns and Fern Allies of South Pacific Islands* (T Nakamura and S
Matsumoto, Eds.). Tokai University Press.

**Nayar BK, Kaur S**. **1971**. Gametophytes of homosporous ferns. *The
Botanical Review* **37**: 295--396.

**Pagel M**. **1994**. Detecting correlated evolution on phylogenies: A
general method for the comparative analysis of discrete characters.
*Proceedings of the Royal Society B: Biological Sciences* **255**:
37--45.

**Palmer DD**. **2003**. *Hawaii's Ferns and Fern Allies*. Honolulu:
University of Hawaii Press.

**Royal Botanic Gardens and Domain Trust**. **2004**. *Flora of New
South Wales. PlantNET (The NSW Plant Information Network System)* [WWW
document] URL <http://plantnet.rbgsyd.nsw.gov.au/floraonline.htm>.
[accessed 1 January 2020].

**Tigerschiöld E**. **1989**. Dehiscence of antheridia in thelypteroid
ferns. *Nordic Journal of Botany* **9**: 407--412.

**Tigerschiöld E**. **1990**. Gametophytes of some Ceylonese species of
Thelypteridaceae. *Nordic Journal of Botany* **9**: 657--664.

**Zhang K, Shi L, Zhang X, Jiang C, Tim-Chun W**. **2008**. Gametophyte
morphology and development of six chinese species of *Pteris*
(Pteridaceae). *American Fern Journal* **98**: 33--41.